### 2021-04-06 README.Rmd
Added readme

---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wagglefit

[![R-CMD-check](https://github.com/joseph-palmer/wagglefit/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/check-standard.yaml)
[![lint](https://github.com/joseph-palmer/wagglefit/actions/workflows/lint.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/lint.yaml)
[![pkgdown](https://github.com/joseph-palmer/wagglefit/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/pkgdown.yaml)
[![test-coverage](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml)

The goal of wagglefit is to allow for easy waggle dance analysis.

## Installation

You can install the released version of wagglefit from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("wagglefit")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("joseph-palmer/wagglefit")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(wagglefit)
## basic example code
```