### 2021-06-08 analysis/analysis.Rmd
created Rmd document for making manuscript figures

# Honeybee foraging analysis
## Joseph Palmer
---

### Synopsis

This document contains the analysis of models fit to the data using `run_all_models()`

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
library(cowplot)
library(dplyr)
library(tibble)
source("analysis/fit_data.R")
```

```{r}

path <- "analysis/results_version1"
data <- get_data()

sites <- purrr::map_chr(
  list.files(path),
  ~ {
    return(gsub(".rds", "", .x))
  }
)

plots_data <- purrr::map(
  sites,
  ~ {
    subdat <- data %>% dplyr::filter(.data$site == .x)
    df <- readRDS(paste0(path, "/", .x, ".rds"))[-3]
    results_tibble <- make_results_tibble(df) %>%
      rowwise() %>%
      mutate(
        ks_p = calc_ks(
          subdat$foraging_distance, c(p, ls, ln, qn, a),
          data_name, TRUE
        ),
        ks_d = calc_ks(
          subdat$foraging_distance, c(p, ls, ln, qn, a),
          data_name, FALSE
        ),
        site = .x
      ) %>%
      select(
        site,
        everything()
      )
    ret <- list(
      site_name = list(
        results_tibble,
        make_full_plot(subdat$foraging_distance, df)
      )
    )
    return(ret)
  }
)
```

```{r}
plots <- plots_data %>%
  purrr::map(
    ~ {
      .x[[1]][[2]]
    }
  )

all_data <- plots_data %>%
  purrr::map(
    ~ {
      .x[[1]][[1]]
    }
  ) %>%
  bind_rows()

aic_plot <- all_data %>%
  group_by(site) %>%
  slice(which.min(AIC)) %>%
  select(data_name) %>%
  group_by(data_name) %>%
  summarise(lowest_AIC = n()) %>%
  ggplot(aes(x = data_name, y = lowest_AIC)) +
  geom_bar(stat = "identity", fill = "#0084ff") +
  labs(x = "Model", y = "Number of lowest AIC sites")
```