### 2021-09-02 analysis/proportions-of-scouts.Rmd

# Analysis of the proportion of scouts across environmental gradients
## Joseph Palmer
---

To start I am just having a play with the proportion of scouts from each model.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(family = "URWHelvetica", size = 8)
    )
)
library(cowplot)
library(dplyr)
library(tibble)
source("analysis/fit_data.R")
```

```{r}

# load analysis results
path <- "analysis/results_version1"
data <- get_data()
sites <- purrr::map_chr(
  list.files(path),
  ~ {
    return(gsub(".rds", "", .x))
  }
)

# read analysis results into a single dataframe
model_results <- purrr::map(
    sites,
    ~ {
      subdat <- data %>% dplyr::filter(.data$site == .x)
      df <- readRDS(paste0(path, "/", .x, ".rds"))[-3]
      results_tibble <- make_results_tibble(df)
      results_tibble$site = .x
      return(results_tibble)
    }
  ) %>%
  bind_rows()

# keep only best fitting model
model_results <- model_results %>% group_by(site) %>% top_n(-1, AIC)

# load raw data
df <- read.table('analysis/data/FullHBForagingData.csv', header = TRUE,  sep = ',',  stringsAsFactors = FALSE)

# filter for landscape type
df_site_info <- df %>% select(site, LU) %>% unique()

# join model results with landscape type
model_results_pscouts <- model_results %>%
  filter(data_name == "all") %>%
  inner_join(df_site_info, by = c("site" = "site"))

# point plot of proportion of scouts for each site, coloured by landscape type
model_results_pscouts %>%
  ggplot(aes(x = site, y = p)) +
  geom_point(aes(color = LU))

```

The preceding code block show the proportion of scouts in each urban and rural site.

Lets look at summary stats for each landscape tyepe.

```{r}

model_results_pscouts %>%
  ggplot(aes(x = LU, y = p)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center")

```

There is a much higher spread in the rural sites, but no real difference in the proportion of scouts between each groups.

Need to work more on this and write more on what we should do with this....

Also need to ask vincent what he thinks of this chapter idea. Send him an overview and some data to look over before our next meeting.

```{r}
# bootstrapped mean / median
boot.median <- function(data, n = 10000) {
  resamples <- purrr::map_dbl(
    seq_len(n),
    ~ {
      median(sample(data$p, replace = TRUE))
    }
  )
  std.err <- sqrt(var(resamples))
  return(list(std.err = std.err, medians = resamples))
}

urban.boot <- model_results_pscouts %>%
  filter(LU == "Urban") %>%
  select(p) %>%
  boot.median()

hist(urban.boot$medians)

rural.boot <- model_results_pscouts %>%
  filter(LU == "Rural") %>%
  select(p) %>%
  boot.median()

hist(rural.boot$medians)

```

```{r}

# beta regression
library(betareg)

model_results_pscouts

betamod <- betareg(p ~ LU, data = model_results_pscouts)
summary(betamod)
model_results_pscouts$predbeta = predict(betamod, interval = "confidence")

ggplot(data = model_results_pscouts, aes(x = LU, y = p)) +
  geom_point() +
  geom_point(aes(y = predbeta, color = LU))

```

The above plot (code to make the beta plot) fits a beta regression. The beta regression is to be used when the predictor variables are proprtions. Whilst there is no difference between the proportion of scouts in urban and rural landscapes, there is a much tighter spread in the urban colonies than the rural colonies.