### 2021-10-05 analysis/analyse_proportion_of_scouts.Rmd
reorganised the code structures and stated exploring how the proportion of scouts varies accross landscapes.

# Analysing the proportion of scouts estimated in different environments
## Joseph Palmer
---

```{r, include = FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 8)
    )
)
```

First we load in the data and make the tables required. I.e. site land type information along with the estimates proportion of scouts from the model fits in `fit_models.Rmd`.

```{r}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("analysis/results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "analysis/data/urban-landuse.csv")
rural_lu <- read.csv(file = "analysis/data/agri-rural-landuse.csv")

# put proportion of scouts into landuse data
urban_lu_long <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu) %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu_long <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu) %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

Now we have everything in, lests do a basic plotting to see what the relationships between each landuse type and the proportion of scouts looks like.

For rural we can see that for most catagories there isn't much of a relationship, however, from a basic linear model the proportion of scouts seems to increase with arable and decrease with more built up areas.

```{r}

rural_lu_long %>%
  ggplot(aes(x = p, y = values)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~landuse, ncol = 2) +
  theme(
    text = element_text(size = 8)
  )
```

for urban we don't really see much at all. There is a slight increase the proportion of scouts in a continous central landscape, and a decrese in sparse residential areas. Everything else seems rather flat.

```{r}

urban_lu_long %>%
  ggplot(aes(x = p, y = values)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~landuse, ncol = 2) +
  theme(
    text = element_text(size = 8)
  )
```