### 2021-10-05 analysis/analyse_proportion_of_scouts.Rmd
started with the PCA analysis, see if it makes sense

# Analysing the proportion of scouts estimated in different environments
## Joseph Palmer
---

## Set up data and environment

```{r, include = FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 8)
    )
)
```

First we load in the data and make the tables required. I.e. site land type information along with the estimates proportion of scouts from the model fits in `fit_models.Rmd`.

```{r}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("analysis/results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "analysis/data/urban-landuse.csv")
rural_lu <- read.csv(file = "analysis/data/agri-rural-landuse.csv")

# put proportion of scouts into landuse data
urban_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu)

urban_lu_long <- urban_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu)
rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

Now we have everything in, lests do a basic plotting to see what the relationships between each landuse type and the proportion of scouts looks like.

## Basic analysis

For rural we can see that for most catagories there isn't much of a relationship, however, from a basic linear model the proportion of scouts seems to increase with arable and decrease with more built up areas.

```{r}

rural_lu_long %>%
  ggplot(aes(x = p, y = values)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~landuse, ncol = 2) +
  theme(
    text = element_text(size = 8)
  )
```

for urban we don't really see much at all. There is a slight increase the proportion of scouts in a continous central landscape, and a decrese in sparse residential areas. Everything else seems rather flat.

```{r}

urban_lu_long %>%
  ggplot(aes(x = p, y = values)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~landuse, ncol = 2) +
  theme(
    text = element_text(size = 8)
  )

```

## Princple Component Analysis

```{r}

prcomp(urban_lu[, 3:length(names(urban_lu))])
```