### 2022-08-26 README.Rmd
added DOI badge to readme

---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wagglefit

[![R-CMD-check](https://github.com/joseph-palmer/wagglefit/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/check-standard.yaml)
[![lint](https://github.com/joseph-palmer/wagglefit/actions/workflows/lint.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/lint.yaml)
[![pkgdown](https://github.com/joseph-palmer/wagglefit/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/pkgdown.yaml)
[![test-coverage](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/joseph-palmer/wagglefit/actions/workflows/test-coverage.yaml)
[![DOI](https://zenodo.org/badge/353413244.svg)](https://zenodo.org/badge/latestdoi/353413244)

wagglefit is an R package to analyse waggle run distributions and estimate the resulting proportions of scout and recruit dances, providing an estimate of dance use in undisturbed colonies.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("joseph-palmer/wagglefit")
```

## Honeybees vary collective decision making in different landscapes

All code used in the publication *Honeybees vary collective decision making in different landscapes* is stored in `inst/publication.tar.gz/`. To extract it use `tar -xzvf publication.tar.gz`.

## Reporting issues

If you notice any issues with this code please raise an issue in the github repo.